High-Performance Data Analysis on Janus using Apache Spark

### 2013-12-17 paper.tex
added efficiency, karp-flatt, big-grep results, and grep discussion

\section{Evaluation}
We conducted our evaluation of the Spark framework on the University of
Colorado's Janus supercomputer.  Janus consists of 1,368 nodes, each containing
two six core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total of 12 cores
per node.  Additionally, each core has 2 GB of 1333 MHz DDR RAM, for a total of
24 GB per node \cite{JanusPaper}. In our research, we employed Spark version
0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
we access input files from Janus' **lustre distrubted file system**. % WHAT ELSE
Our aim was to assess the Spark on Janus and determine optimal configurations
to achieve the best performance.

\subsection{Testing}
Following the example from Google's first paper on MapReduce, the first test we
performed was a parallel grep. Here an application would scan through a large
test file, looking for lines that matched a given regular expression, finally
returning the total number of matched lines. Two series of test were run on two
different files sizes. We first wanted to see how Spark scales in an ideal
setting, where the entire data set fits in memory. ***This way we avoid the
large performance hit Spark takes writing intermediate results to disk***.
Given that each node has a total of 24 GB of RAM, we limited our first problem
size to 20GB. This allowed a single node to contain the entire problem in
memory, leaving some space for the OS and other supporting processes. We
excuted this first test with 1 to 5 nodes. The second test we performed was
run with 5 to 50 nodes, each processing a 100GB file.

@@

% in graphRef we can see how well the cores are being utilized
% in an efficiency graph
% as expected from previous graphs, efficiency not great
% starts under 40% efficiency with 1 node
% end up at just about 10% efficiency with 5 nodes, or 60 cores

@@

% finally we see this data in terms of karp-flatt metric in graphRef
% with the exception of the 12 workers per node in pink
% we see that the karp-flatt metric remains fairly constaint
% values between 0.12 and 0.14
% this constant value would suggest that the serial runtime
% is what is preventing better scaling

@@

% to see if we could see better scaling over a largest file
% switch from 20GB to 100GB file
% accessed file from lustre drive
% as 100GB cannot fit in memory of one node, start with 5
% graphRef should the runtimes over this large file
% clearly there appears to be no change in runtimes
% remains round 500 seconds between 5 nodes and 50
% suspecting short compute time of grep expression
% switch to a more computational expensive and iterative PageRank

@@

\section{Discussion}
% the results for the grep test over the 20gb file
% gave us some insight on how well such a problem scale in spark
% for that given task and problem size
% 3 nodes appears to be a good balance between speedup and efficiency
% we all saw how workers per node effects runtime
% while 1-3 workers remains close, clear 6-12 is not ideal
% it seems as if one my observer slightly better results
% with 2-3 nodes over 1 node, but not by much

% in the large grep test over the 100 gb file
% suprisingly obvserved no change in performance
% perhaps the results of karp-flatt metric in previous test
% gives us same insight as to why we are not seeing improvement
% we suspect the long load times are are bigger
% than the relatively short time needed to graph the loaded information
% the long load times may suggest the file is not read in parallel

>>>>

\section{Evaluation}
We conducted our evaluation of the Spark framework on the University of
Colorado's Janus supercomputer.  Janus consists of 1,368 nodes, each containing
two six core Intel Xeon Westmere-EP chips at 2.8 GHz, for a total of 12 cores
per node.  Additionally, each core has 2 GB of 1333 MHz DDR RAM, for a total of
24 GB per node \citep{tufo}. In our research, we employed Spark version
0.8.0 along with Scala version 2.9.3. During our testing of the Spark framework,
we access input files from Janus' **lustre distributed file system**. % WHAT ELSE
Our aim was to assess the Spark on Janus and determine optimal configurations
to achieve the best performance.

\subsection{Testing}
Following the example from Google's first paper on MapReduce, the first test we
performed was a parallel grep \citep{dean-mapreduce}. Here an application would
scan through a large test file, looking for lines that matched a given regular
expression, finally returning the total number of matched lines. Two series of
test were run on two different files sizes. We first wanted to see how Spark
scales in an ideal setting, where the entire data set fits in memory. ***This
way we avoid the large performance hit Spark takes writing intermediate results
to disk***.  Given that each node has a total of 24 GB of RAM, we limited our
first problem size to 20GB. This allowed a single node to contain the entire
problem in memory, leaving some space for the OS and other supporting
processes. We executed this first test with 1 to 5 nodes. The second test we
performed was run with 5 to 50 nodes, each processing a 100GB file.

@@

In the efficiency graph show in Figure \ref{fig:workNodeEff} we can see just
how well the cores are being utilized. As expected from the results shown in
the previous figures, the efficiency is quite poor. Employing a single node
starts just under 40\% efficiency. By the time we scale the problem to 5 nodes
(60 cores) we are just above 10\% efficiency.

@@

Finally, Figure \ref{fig:workNodeKF} reflects the Karp-Flatt metric for the
runtime results. With the exception of the 12 workers per node (pink line), we
we see that the Karp-Flatt metric remains relatively constant, with values
between 0.12 and 0.14.  This constraint value would suggest that the large serial
portion of the application is what is preventing use from seeing better
scaling.

@@

To see if we could observe better scaling, we ran the same grep test over a
100 GB file.  During this test, we also increased the number of nodes used,
starting with 5 and going to 50 nodes. Given its size, the input text file was
accessed from the Janus' lustre filesystem. Figure \ref{fig:bigTime} shows the
runtimes observed over this large problem.  Clearly there appears to be no
change in runtimes, as the runtimes hover around 500 seconds, regardless of the
number of nodes used. Suspecting the short computational time of the grep
search, we began running our tests on the more computationally expensive
PageRank algorithm.

@@

\section{Discussion}
The results collected from the initial grep test over the 20 GB file gave us
some insight on how well such a problem scales in the Spark framework.  Given
that text file of 20 GB, it appears that employing 3 nodes provides a good
balance between speedup and efficiency.  We also saw how the number of workers
per node effects the overall runtime. While 1 to 3 workers yield relatively
similar results, it seems that using 2 to 3 nodes may give a slight performance
increase. This test would need to be run several more times to achieve more
more conclusive runtime averages.  It would also be beneficial to increase the
input file size and node codes, to see if the observed trend continues as the
problem scales.

In the grep test over the 100 GB file, surprisingly we observed practically no
change in runtime performance. Perhaps the Karp-Flatt metric results in the
previous test gives us some insight as to why this is the case. We suspect the
long load times on this larger file greatly outweigh any speedup achieved
through running grep on multiple workers. Given that, we also suspect that
Spark could be reading this file in serial from the lustre file system. The
lack of Spark logging and profiling tools makes it difficult to determine
exactly how long each stage is takes. But further investigation would be needed
to determine a possible solution or workaround.
